How did the introduction of Stemmed Points affect Mobility and Site Occupation during the Late Pleistocene in Korea?

### 2021-04-14 analysis/paper/paper.Rmd
knit ok with SPD figure

Figure \@ref(fig:volume-density-retouch) shows a strong pattern of artifacts from less dense assemblages having higher retouch frequencies while more dense assemblages have lower retouch frequencies (`r apa::apa(den_sp_ttest)`), further showing a spectrum of site functions in this sample. As we saw for artifact density, there are no clear directional chronological trends in retouch frequencies. Assemblages containing stemmed points tend to have substantially fewer retouched pieces compared to assemblages without stemmed points (`r apa::apa(retouch_sp_ttest)`).

@@

Figure \@ref(fig:site-elevation) shows that people occupied sites located at similar elevations, around 0 to 200 meters above sea level. As we saw for artifact density, there is an increase in the upper bound of site elevation in more recent times. The most striking pattern in the elevation data is the substantial difference in elevation between sites with stemmed points and sites without. The distribution of elevations of sites without stemmed points is generally lower than the elevations of sites with stemmed points. Although not a statistically significant difference in elevations (`r apa::apa(elev_sp_ttest)`), this may indicate that forager groups who used stemmed points were generally more able to occupy higher altitudes, while the groups without stemmed points tended to prefer lower altitudes. According to our locally weighted regression, sites without stemmed points reach a maximum elevation of about 150 m at around 39 ka. Sites containing stemmed points reach a maximum elevation of 175 m at 28 ka.

@@

```{r spd-model, fig.cap=paste0("Summed probability distribution of ", nrow(dates_clean), " radiocarbon dates. The black solid line represents actual radiocarbon ages. The grey shaded region shows the Monte Carlo envelope that encompasses the 95% confidence interval for the null models. The red and blue vertical bands highlight the portions of the SPD where positive and negative deviations are detected. We applied logistic, exponential and linear to AIC along with the logistical growth model (fit) to evaluate each model [@riris2019widespread]. We excluded the uniform model because it has no fitted curve or surface to evaluate. The best fit model has the lowest AIC score. In this case, the linear model is the best fit for the observed data"), cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/009-summed-probability-distributions.R"))}
quietly(source(here::here("analysis/paper/009-summed-probability-distributions.R")))
knitr::include_graphics(here::here("analysis/figures/008-summed-probability-distribution-models.png"))
```

>>>>

Figure \@ref(fig:volume-density-retouch) shows a strong pattern of artifacts from less dense assemblages having higher retouch frequencies while more dense assemblages have lower retouch frequencies (`r den_sp_ttest_str`), further showing a spectrum of site functions in this sample. As we saw for artifact density, there are no clear directional chronological trends in retouch frequencies. Assemblages containing stemmed points tend to have substantially fewer retouched pieces compared to assemblages without stemmed points (`r retouch_sp_ttest_str`).

@@

Figure \@ref(fig:site-elevation) shows that people occupied sites located at similar elevations, around 0 to 200 meters above sea level. As we saw for artifact density, there is an increase in the upper bound of site elevation in more recent times. The most striking pattern in the elevation data is the substantial difference in elevation between sites with stemmed points and sites without. The distribution of elevations of sites without stemmed points is generally lower than the elevations of sites with stemmed points. Although not a statistically significant difference in elevations (`r elev_sp_ttest_str`), this may indicate that forager groups who used stemmed points were generally more able to occupy higher altitudes, while the groups without stemmed points tended to prefer lower altitudes. According to our locally weighted regression, sites without stemmed points reach a maximum elevation of about 150 m at around 39 ka. Sites containing stemmed points reach a maximum elevation of 175 m at 28 ka.

@@

```{r spd-model, fig.cap=paste0("Summed probability distribution of ", nrow(dates_clean), " radiocarbon dates. The black solid line represents actual radiocarbon ages. The grey shaded region shows the Monte Carlo envelope that encompasses the 95% confidence interval for the null models. The red and blue vertical bands highlight the portions of the SPD where positive and negative deviations are detected. We applied logistic, exponential and linear to AIC along with the logistical growth model (fit) to evaluate each model [@riris2019widespread]. We excluded the uniform model because it has no fitted curve or surface to evaluate. The best fit model has the lowest AIC score. In this case, the linear model is the best fit for the observed data"), cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/009-summed-probability-distributions.R"))}
# quietly(source(here::here("analysis/paper/009-summed-probability-distributions.R")))
knitr::include_graphics(here::here("analysis/figures/008-summed-probability-distribution-models.png"))
```
